A formal approach but practical approach in train regulation

### 2025-06-13 paper/main.tex
merge
des todo en fin de section 2

Now, we come to the train model. Each train has an identifier. In sidother words, we have a set of trains $\trains = \{1, \ldots, k\}$. Each train has a position on the network. The position is the index of the block it stays in or the last block it was in. Next, a train has a direction in $\directions$. The direction does not mean the train is running---it can be blocked by some red signal---but that its next move will be in that direction. Finally, a train has orders. Orders are given as a sequence of orders following the grammar:
$$  seq ::= \su(dir,pos) $$
with $dir \in \directions$ and $pos \in \blocks$. An order $\su(dir, pos)$ states that the train direction is set to $dir$ and that the train runs until it reaches $pos$.  The sequence may be empty.

@@

The regulator is characterized by 6 fields. First, it has a general program, next mentioned as $E$. The program contains the plan of the regulator. The regulator will follow its rule. A program is given as a sequence of statements. Those follow the grammar:

\begin{grammar}[Regulator's rules][h][regulateur]
    \firstcasesubtil{stmt}{ev(id, n) : atoms}{}
    \firstcasesubtil{atoms}{\incr(j)}{}
    \otherform{att(j, v)}{}
    \otherform{\turn(sid, s)}{}
    \otherform{auth}{}
    \otherform{atoms ; atoms}{}
    \otherform{\varepsilon}{}
\end{grammar}
with $id \in \trains$, $n \in  \Nat$, $j \in \blocks$, $sid \in \turnouts$, $s\in \internalState$.  All rules are guarded by a condition $ev(id,n)$ stating that train $id$ reached the $n$-th step of its route. The atoms update the regulator fields. Before we describe them and the other fields, a notation: given some function $f: X \to Y$, $x \in X$ and $y \in Y$, we denote by $f[x \leftarrow y]$, the function:
 $$f[x \leftarrow y] \triangleq \left\{  \begin{array}{ll} x &\mapsto y\\
x' \neq x \in X &\mapsto f(x')
\end{array}\right.$$

@@

The waiting field is a function $W: \blocks \times \Nat \to \trains \times Bool$. XXX

The buffer field is the XXX

The history field $H: \trains \to \blocks$ gives the last position of each train recorded  by the regulator.

So, the regulator is described as a tuple $R = (E, T, W, B, H)$. Applying $\incr(j)$ to the regulator is applying it to the token field: $R \cdot \incr(j) = (E, T \cdot \incr(j), W, B, H)$. We use the same convention for the other atoms.

@@

>>>>

Now, we come to the train model. Each train has an identifier. Without loss of generality, the identifiers form a set  $\trains = \{1, \ldots, k\}$. Each train has a position on the network. The position is the index of the block it stays in or the last block it left. Next, a train has a direction in $\directions$. The direction does not mean the train is running---it can be blocked by some red signal---but that its next move will be in that direction. Finally, a train has orders. Orders are given as a sequence of orders following the grammar:
$$  seq ::= \su(dir,pos) $$
with $dir \in \directions$ and $pos \in \blocks$. An order $\su(d, pid)$ states that the train direction is set to $d$ and that the train runs until it reaches $pid$.  The sequence may be empty.

@@

To help the reader, in the sequel, by convention, we use $pid$ for blocks, $sid$ for trackworks identifiers (switch id) and $tid$ for train identifiers.

The regulator is characterized by 5 fields. First, it has a general program, next mentioned as $E$. The program is the plan of the regulator. The regulator will follow its rules. A program is given as a sequence of statements. Those follow the grammar:
\begin{grammar}[Regulator's rules][h][regulateur]
    \firstcasesubtil{stmt}{ev(tid, n) : (atom)*}{}
    \firstcasesubtil{atom}{\incr(pid)}{}
    \otherform{att(pid, v)}{}
    \otherform{\turn(sid, s)}{}
    \otherform{auth}{}
 \end{grammar}
with $tid \in \trains$, $n \in  \Nat$, $pid \in \blocks$, $sid \in \turnouts$, $s\in \internalState$.  All statements are guarded by a condition $ev(tid,n)$ stating that train $tid$ reached the $n$-th step of its route. The atoms update the regulator fields. Before we describe the intention of atoms and the other fields, a notation: given some function $f: X \to Y$, $x \in X$ and $y \in Y$, we denote by $f[x \leftarrow y]$, the function:
 $$f[x \leftarrow y] \triangleq \left\{  \begin{array}{ll} x &\mapsto y\\
x' \neq x \in X &\mapsto f(x')
\end{array}\right.$$

@@

The waiting field is a function $W: \blocks \times \Nat \to \trains \times Bool$. It states that block must be protected. \todo{donner plus de précision sur les intentions ici.}

The history field $H: \trains \to \blocks$ gives the last position of each train recorded  by the regulator. \todo{A quoi cela sert il en bref.}

So, the regulator is described as a tuple $R = (E, T, S, W, H)$. Applying $\incr(j)$ to the regulator is applying it to the token field: $R \cdot \incr(j) = (E, T \cdot \incr(j), S, W, B, H)$. We use the same convention for the other atoms.

\todo{Je pense que le régulateur ne devrait pas contenir plus d'éléments. Les autres sont en fait des meta-données qui gèrent "le baton de parole" (les priorités sur les règles). Donc, G et B devraient être en dehors du tuple.}

@@

Description par des règles. La gestion du temps est "discrétisée". On ne peut pas "choisir" quelle est la bonne règle. Gestion par des gardes, et des messages. Il y a des soucis grand pas, petit pas
